### 2021-02-12 manuscript.Rmd
initial version of OCRbayes

---
output: word_document
bibliography: /home/xiang/Documents/Bibliography/seahorse.bib
csl: plos-one.csl
---

OCRbayes: A Bayesian hierarchical modeling framework for Seahorse extracellular flux oxygen consumption rate data analysis

Xiang Zhang, 
Taolin Yuan, 
Keijer Jaap,
Vincent C. J. de Boer

Human and Animal Physiology, Wageningen University, De Elst 1, 6708 WD Wageningen, The Netherlands

# Abstract

Mitochondrial dysfunction is involved in many complex diseases. 
Efficient and accurate evaluation of mitochondrial functionality is crucial for understanding pathology as well as facilitating novel therapeutic developments. 
As a popular platform, Seahorse extracellular flux (XF) analyzer is widely used for measuring mitochondrial oxygen consumption rate (OCR) in living cells. 
A hidden feature of Seahorse XF OCR data is that it has a complex data structure, caused by nesting and crossing between measurement cycles, wells and plates. 
Surprisingly, statistical analysis of Seahorse XF data has not received sufficient attention, and current methods completely ignore the complex data structure, impairing the robustness of statistical inference. 
To rigorously incorporate the complex structure into data analysis, here we developed a Bayesian hierarchical modeling framework, OCRbayes, and demonstrated its applicability based on analysis of published data sets. 
We showed that OCRbayes can analyze seahorse XF OCR data with either single or multiple plates, and it has potential to be used for diagnosing patients with mitochondrial diseases.

# Introduction

Mitochondria are double-membrane organelles that are central hubs in regulating energy generation and partitioning. 
Since defects in mitochondrial function have been found to be involved in cardiovascular diseases [@Ballinger2005], type II diabetes [@Szendroedi2011], neurodegenerative disease [@Johri2012] and cancer [@Erez2015], restoring mitochondrial function is emerging to be a therapeutic target for common diseases [@Murphy2018].

Mitochondria produce energy primarily by transferring electrons along the electron transport chain (ETC) on the inner membrane. 
Along the ETC, there are four complexes (including complex I, II, III and IV), building up a proton gradient that is ultimately used by ATP synthase to generate ATP. 
Since electrons are primarily accepted by O~2~ to produce H~2~O, the activity of ETC can be assessed by measuring oxygen consumption rate (OCR). 
As a reliable and efficient platform, Seahorse XF analyzer is a multiwell plate based respirometry assay that is widely used to quantify OCR in living cells [@Wu2007]. 
Typically, the seahorse XF analyzer measures the OCR of cells in a 96-well plate under different ETC perturbation scenarios, used for assessing mitochondrial functionality such as maximal respiration, leak respiration and ATP-linked respiration [@@Chacko2014]. 
Seahorse XF Cell Mito Stress Test consists of four phases including an initial phase, followed by three consecutive injections, including oligomycin, FCCP (carbonyl cyanide-4-(trifluoromethoxy)phenylhydrazone) and a mixture of rotenone and antimycin A. 
In the initial phase, no compound is injected and ETC is not perturbed. 
Once oligomycin is injected in the assay, proton translocation through ATP synthase is blocked, leading to a reduction in OCR. 
Next, injection of FCCP allows protons to move into the mitochondrial matrix independent of the ATP synthase, pushing the oxygen consumption to the maximum level. 
In the end, a mixture of rotenone and antimycin A is injected to inhibit complex I and III, shutting down mitochondrial respiration.

A standard seahorse run includes three measurement cycles for each phase. 
Every measurement cycle starts by lowering the cartridge and creating a temporary semi-closed 7 $\mu$L chamber [@Gerencser2009]. 
During a measurement cycle, fluorescent oxygen sensors capture oxygen concentration changes in the chamber and outputs OCR. 
The Seahorse XF analyzer measures OCR in tens to hundreds of thousands of cells per well and typically requires 4–5 replicate wells per experimental group [@Artyomov2020].
A single 96-well plate can contain multiple experimental groups. 
Since measurement cycles are nested within phases, phases are crossed with wells, and wells are nested in experimental groups, a very complex structure is inherently embedded in the Seahorse XF OCR data. 
Surprisingly, current Seahorse XF data analyses most often ignore this complex structure, and by default, data sets are often chopped into subgroups followed by performing ANOVA-like statistical tests. 
Although there are tools developed for seahorse data analysis, none of them takes the complexity of the data structure into account. 
In addition, these methods differ in how to extract OCR value for each phase. 
For instance, OCR-Stats [@Yepez2018] took OCR value from the last measurement cycle in phase 1, the minimal OCR value in phase 2, the maximal OCR value in phase 3 and the minimal OCR value in phase 4. In contrast, SHORE [@Nicholas2017] took the mean OCR value in phase 1, the median OCR value in phase 2, the maximal OCR value in phase 3, and the median OCR value in phase 4.

A natural way to incorporate the complex structure into data analysis is to use hierarchical modeling, which has been extensively developed for analyzing gene expression data  [@Robinson2010a;@Love2014a]. 
Here we developed a Bayesian hierarchical modeling framework, OCRbayes, for the Seahorse XF OCR data analysis. 
OCRbayes is based on a fully Bayesian approach due to its flexibility and convenience for constructing the hierarchical models. 
We analyzed a published OCR data set OCR data set to demonstrate the applicability of our approach analyzing Seahorse XF OCR data as well as its potential usage for mitochondrial disease diagnostics.

# Method

To demonstrate how OCRbayes can be used for Seahorse XF OCR data analysis, we analyzed 26 human fibroblasts that were derived from patients that were suspected to have mitochondrial dysfunction. 
In particular, six of these cell lines were derived from patients with rare mutations in genes associated with an established cellular respiratory defect, and they serve as positive controls. 
Meanwhile, there are two cell lines that showed no respiratory defects in earlier studies [@Powell2015;@Kremer2016] and they served as negative controls. 
Since fold change of maximal respiration is the most common metric used for identifying mitochondrial dysfunction, we compared the maximal respiration of all these 26 cell lines against the control cell line.

## Data

We downloaded the seahorse XF OCR data presented in OCR-Stats [@Yepez2018] from the journal website https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0199938#sec022. 
This data set contains 203 human fibroblasts assayed in 126 plates. 
The normal human dermal fibroblasts (NHDF) is used as control. 
The rest 202 cell lines were derived from patients suspected to suffer from rare mitochondrial diseases. 
Among these 202 cell lines, 26 fibroblast cell lines were grown independently and were measured in multiple plates. 
The rest 176 fibroblast cell lines were measured in a single plate. 
We then spitted the data set into two subsets. 
One subset, which contains the 176 fibroblasts that were assayed in a single plate, was used for estimating prior distributions. 
The other subset, which contains the 26 fibroblasts, was used for benchmarking.

Since dealing missing observations is not the focus of this study, we removed the entire measurement values of a well if it contains a missing OCR value. 
In addition, since we focused on the standard mitochondrial stress test assay, we removed any well if that well does not have 4 phases with 3 measurement cycles each. 
For the prior distribution subset, we had 177 (including the control cell line NHDF) cell lines on 78 plates after filtering. 
Since NHDF cell line was present in all 78 plates, the OCR values derived from NHDF cell line were used for estimating the prior distribution of between plate variation.

## Implementation details

The Bayesian hierarchical models were implemented in Stan (version 2.19.3) [@Carpenter2017]. 
We fitted the model by running Hamiltonian Markov Chain Monte Carlo. 
We ran four Markov chains with 2000 iterations in each chain. 
The code can be found at https://github.com/XiangZhangSC/seahorse. 
We also provided the R code of OCRbayes in the supplementary file.

## Calculation of bioenergetic measures

Based on OCR (per 1000 cells), we calculated various bioenergetic measures, such as basal respiration, proton leak, ATP-linked respiration, spare respiratory capacity and maximal respiration.
These bioenergetic measures are defined as below.

$$\text{Basal respiration} = \text{OCR}_\text{per 1k cells, initial} - \text{OCR}_\text{per 1k cells, rotenone} \text{ (Equation 1)}$$
$$\text{ATP-linked respiration} = \text{OCR}_\text{per 1k cells, initial} - \text{OCR}_\text{per 1k cells, oligomycin} \text{ (Equation 2)}$$
$$\text{Proton leak} = \text{OCR}_\text{per 1k cells, oligomycin} - \text{OCR}_\text{per 1k cells, rotenone} \text{ (Equation 3)}$$
$$\text{Spare respiratory capacity} = \text{OCR}_\text{per 1k cells, FCCP} - \text{OCR}_\text{per 1k cells, initial} \text{ (Equation 4)}$$
$$\text{Maximal respiration} = \text{OCR}_\text{per 1k cells, FCCP} - \text{OCR}_\text{per 1k cells, rotenone} \text{ (Equation 5)}$$

## Statistical analysis

For each cell line study, we compared the patient cell line to the control cell line (NHDF). 
Since maximal respiration is the primary readout of mitochondrial defect in literature, and was also evaluated in the OCR-Stats, here we reported the mean log2 fold change of maximal respiration together with the False Discovery Rate (FDR). 
FDR was calculated based on the posterior error probability ($1 - P(log2(\frac{\text{patient}}{\text{control}}|\text{data}) < 0)$), where $P$ represents the posterior distribution after fitting the Bayesian model to the experimental OCR data. 
If the FDR below 0.05, we consider that difference between patient and control cell line was statistically significant.

# Results

## OCRbayes: A Bayesian hierarchical modeling framework for Seahorse OCR data analysis

In order to incorporate the complex data structure into the analysis of seahorse XF OCR measurements, we developed a Bayesian hierarchical modeling framework, OCRbayes. 
In this study, we focused on experimental data containing two groups, patient and control. 
The mathematical description of our Bayesian hierarchical modeling framework is shown below:

$$\text{OCR}_\text{obs}\text{[i]} \sim \text{Lognormal(logOCR}_\text{true}\text{[P[i],W[i],I[i]]}, \text{logOCR}_\text{se}\text{[P[i],I[i]])} \text{ (Equation 6)}$$
$$\text{logOCR}_\text{true}\text{[P[i],W[i],I[i]]} \sim \text{Normal(}\text{OCR}_\text{per 1k cells}\text{[P[i],G[i],I[i]]} \times \text{N}_\text{cells}\text{[W[i]]}, \sigma_\text{well}\text{[P[i],G[i],I[i]])} \text{ (Equation 7)}$$
$$\text{OCR}_\text{per 1k cells}\text{[P[i],G[i],I[i]]} \sim \text{Lognormal(}\mu_{\text{OCR}_\text{per 1k cells}}\text{[G[i], I[i]]}, \sigma_\text{plate}\text{[G[i], I[i]])} \text{ (Equation 8)}$$

From our perspective, there are three sources of technical variations including 1) between measurement cycle variation, 2) between well variation and 3) between plate variation in the Seahorse XF OCR data.

The between measurement cycle variation refers to the variability between the three OCR values generated by the three measurement cycles within an injection phase. 
In Equation 6, we used a lognormal distribution with the true OCR value at the log scale ($\text{logOCR}_\text{true}$) and the between measurement cycle standard deviation (also at the log scale, $\text{logOCR}_\text{se}$) to model each observed OCR value ($\text{OCR}_\text{obs}$). 
$\text{OCR}_\text{obs}$ is a vector of length $\text{N}_\text{plates} \times \text{N}_\text{wells} \times \text{N}_\text{injections} \times \text{N}_\text{Measurement}$, where $\text{N}_\text{plates}$, $\text{N}_\text{wells}$, $\text{N}_\text{injections}$ and $\text{N}_\text{Measurement}$ are number of plates, wells, injections and measurement cycles, respectively. 
The $\text{logOCR}_\text{true}$ refers to $\text{N}_\text{plates}$ layers of matrices. 
Each layer is a matrix with $\text{N}_\text{wells}$ rows and $\text{N}_\text{injections}$ columns. 
$\text{logOCR}_\text{se}$ is a matrix of $\text{N}_\text{plates}$ rows and $\text{N}_\text{injections}$ columns.

In a typical seahorse XF assay, the same cell line from the same experimental group is often seeded in multiple wells. 
However the number of cells in these replicate wells are not identical, thus causing variation in OCR values between these replicate wells. 
To adjust for the effect of cell number difference, in Equation 7 we modeled the true OCR value as a function of the cell number, and from there estimated $\text{OCR}_\text{per 1k cells}$, which represents OCR value per 1000 cells that received the same treatment (or from the same group) and injection.
$\text{OCR}_\text{per 1k cells}$ is a three dimensional matrix with $\text{N}_\text{Groups}$ layers, and each layer has $\text{N}_\text{Groups}$ rows and $\text{N}_\text{injections}$ columns. 
$\text{N}_\text{cells}$ is a vector of length $\text{N}_\text{wells}$, and every entry represents the cell number in that well. 
In addition to cell number difference, initial conditions, treatment concentration, or fluorophore sleeve calibration can also contribute to between well variation. 
Since these factors were difficult to be measured experimentally, we can only estimate an overall effect based on the residual parameter $\sigma_\text{well}$. 
Thus, $\sigma_\text{well}$ represents the well-to-well variation after accounting for cell number difference.

Last but not the least, the seahorse XF assay for a patient cell line is often repeated on different days and on more than one plates. 
Due to differences in temperature, seeding time, growth time, growth medium, sensor cartridge as well as treatment efficiency, the OCR outcomes will differ between plates. 
To take into account the between plate variation, in Equation 8 we used another lognormal distribution with the true OCR value per 1000 cells at log scale ($\mu_{\text{OCR}_\text{per 1k cells}}$) and the between plate standard deviation ($\sigma_\text{plate}$) to model the OCR value per 1000 cells $\text{OCR}_\text{per 1k cells}$.

In this study, since our case studies focused on human fibroblast cells, we used informative prior distributions for $\text{logOCR}_\text{se}$, $\sigma_\text{well}$, $\sigma_\text{plate}$ and $\mu_{\text{OCR}_\text{per 1k cells}}$. 
Since values of all these parameters must be positive, we used four lognormal distributions. 
We estimated the lognormal distribution parameters based on the Seahorse XF OCR data in OCR-stats [@Yepez2018] by running the fitdistr function built in the MASS r package. 
In this study, our prior distributions are
$$
\begin{align}
\text{logOCR}_\text{se} & \sim \text{Lognormal(-3.23, 0.79)} \\
\sigma_\text{well} & \sim \text{Lognormal(-1.62, 0.53)} \\
\sigma_\text{plate} & \sim \text{Lognormal(-1.18, 0.05)} \\
\mu_{\text{OCR}_\text{per 1k cells}} & \sim \text{Lognormal(0.3, 0.79)}
\end{align}
$$

## OCRbayes uses OCR measurements as inputs and outputs posterior distributions for respiration metrics

OCRbayes works with the raw OCR data, and estimates the OCR per 1000 cells during initial phase (Int1), oligomycin phase (Int2), FCCP phase (Int3) and rotenone phase (Int4) (Figure 1A and 1B). 
Afterwards, the posterior OCR per 1000 cells values were transformed into various respiration metrics, such as ATP-linked respiration, basal respiration, maximal respiration, proton leak and spare respiratory capacity (Figure 1C). 
As an example, here we showed patient 65126 and the control cell line (NHDF) on the corresponding plates. 
The posterior distributions showed the range of values for a respiration parameter. 
The spread of the posterior distributions represented the uncertainty, which is resulted from technical variations that will be discussed in the next section.

![Figure 1 (A) Raw OCR values for patient 65126 and control cell line in two plates. (B) Posterior distributions for OCR per 1000 cells in patient 65126 and control cell line during initial phase (Int1), oligomycin phase (Int2), FCCP phase (Int3) and rotenone phase (Int4). (C) Posterior distributions for respiration metrics in patient 65126 and control cell line.](./posterior_respiration_examples.jpeg)

## OCRbayes accounts for various technical variations during seahorse XF OCR data analysis

To visualize the technical variations, we used the cell line 76065 as an example. 
This cell line was measured on five different plates, allowing us to visualize all three technical variations.  
We observed that in the initial phase (Int1) and FCCP phase (Int3), the between plate variation is dominant source of technical variation, followed by between well and between measurement cycle variations. 
However, in the oligomycin phase (Int2) and rotenone phase (Int4), the between well variation can be as important as the between plate variation.

![Figure 2 Technical variations in seahorse XF OCR data of experiments based on cell line 76065 and control cell line. Int1, Int2, Int3 and Int4 represent initial, oligomycin, FCCP and rotenone phase. All the variations were on log scale.](./technical_variations_76065.jpeg)

## Benchmark

To demonstrate that OCRbayes works properly, we applied it to analyze a published seahorse XF OCR data set containing 26 patient cell lines as well as a control cell line. 
In a previous study, 6 cell lines were labeled as positive controls and 2 cell lines were labeled as negative controls. 
Based on our analysis, we found six patient cell lines that had lower maximal respiration compared to the control cell line with False Discovery Rate (FDR) below 0.05. 
Five of these six cell lines were actually the positive controls. 
Meanwhile, the two negative controls had FDR above 0.05.

![Figure 3 Maximal respiration change patient vs. control on multiple plates. (A) average log2 fold change (y-axis) of maximal respiration of all cell lines repeated across plates (x-axis) and their respective controls, sorted by the False Discovery Rate (FDR). Red and blue dots represent positive and negative controls, respectively. (B) similar to (A), but depicting FDR. Red dashed line represents FDR = 0.05.](./benchmark.jpeg)

Our analysis successfully recalled  all negative controls and five out of six positive controls. 
Interestingly, our analysis showed that there was no significant reduction of maximal respiration in patient 61818, which was used as a positive control. 
This patient cell line was measured by seahorse XF OCR assay on three independent plates on three different days. 
We analyzed these three repeated experiments separately as well as combined them together. 
We found that only the second experiment showed difference in maximal respiration.

![Figure 4 (A) Raw OCR values for patient cell line 61818 in three repeated plates. (B) Posterior distributions of maximal respiration of patient 61818 and the control cell line in the same plate. The top three panels represent analysis of three repeated experiments separately. The last panel represents the data analysis outcome after combing all three repeated experiments.](./posterior_maximal_respiration_61818.jpeg)

## OCRbayes can facilitate decision making about whether testing on multiple plates is necessary

A frequent question regarding seahorse XF analysis is that whether it is enough to perform the assay on a single plate or do we really need repeat the assay multiple times on different plates. 
To address this question, we used the two positive control cell lines 67375 and 76065 as examples, since in our analysis they were shown to have the maximal and minimal effect size among the positive controls, respectively.

The cell line 67375 was measured on two plates. 
Based on the first assay, our analysis shewed that the posterior probability of patient 67375 having decreasing maximal respiration than the control was 93.0%. 
By repeating the experiment once more, the posterior probability increased to 99.4%. 
In contrast, the cell line 76065 was measured on five different plates. 
Our analysis showed that based on the first assay, the posterior probability of patient 76065 having lower maximal respiration than the control was 40.6%. 
Repeating the assay once increased the posterior probability from 40.6% to 72.0%. 
When the assay was repeated for the third, fourth and fifth time, the corresponding posterior probabilities increased to 78.8%, 87.9% and 89.9%, respectively.

![Figure 5 (A) Posterior probability of decreasing maximal respiration in patient 67375 compared to the control cell line in the scenario with 1 or 2 plates. (B) Posterior probability of decreasing maximal respiration in patient 76065 compared to the control cell line in the scenario with 1, 2, 3, 4 or 5 plates](./posterior_max_resp_67375_76065.jpeg)

# Discussion

Although Seahorse XF analyzer is widely used in bioenergetic profiling, its data analysis has not received sufficient attention. 
A hidden feature of seahorse XF OCR data is that it contains a complex data structure. 
The complex data structure is due to the fact that seahorse measurement cycles are nested within injections, injections are crossed with wells, and wells are nested in plates. 
As far as we know, currently there is no data analysis protocols that take into account this complex data structure, impairing the robustness of seahorse XF OCR data analysis outcomes. 
This is because when one ignores the data structure, one also ignores the variations between measurement cycles, between wells and between plates. 
In order to make the seahorse data analysis more robust, in this study we developed a Bayesian hierarchical modeling approach, OCRbayes, which accounts for all these technical variations during the data analysis.

## Seahorse XF OCR measurements are noisy

An OCR value is determined not only by mitochondrial activity, but also by technical noise including 1) between measurement cycle variation, 2) between well variation and 3) between plate variation.

Every phase contains three measurement cycles, resulting in three OCR values. 
Since every measurement cycle starts with a "mix and wait" steps to ensure the same baseline of cell values, cell physiology should not substantially change within a phase. 
Existing tools such as OCR-stats [@Yepez2018] and SHORE [@Nicholas2017] use different strategies to select a single data point to represent an injection phase. 
When different data points were chosen, one can get different outcomes, making the current seahorse XF data analysis less robust. 
To avoid this ambiguity, our approach did not select any particular data point, instead modeled all three OCR data. 
By doing so, we incorporated the uncertainty about "which data point should I choose?" into the data analysis and focused on the average behavior.

The between well variation refers to a common observation that OCR values differ among the replicate wells.
One important factor leading to the variation between replicate wells is that cell numbers are not identical in these wells. 
The wells with more cells would have higher OCR than the wells containing fewer cells. 
Fortunately, cell number in each well can be quantified experimentally and used for the data analysis [@Little2020; @Janssen2021]. 
In addition to cell number difference, initial conditions, treatment concentration, or fluorophore sleeve calibration can also contribute to variation between wells. 
OCRbayes also takes into account the between well variation caused by these unobserved factors.

Between plate variation takes place when the same seahorse experiment is repeated on different days and on more than one plates. 
Due to differences in temperature, seeding time, growth time, growth medium, sensor cartridge as well as treatment efficiency, the OCR outcomes will differ between plates [@Koopman2016; @Yepez2018]. 
Often the between plate variation is assumed to be the dominant technical variation involved in seahorse OCR data. 
Based on our analysis, we showed that this assumption may be appropriate for OCR measurements derived from the initial and FCCP phase, but may not work for the OCR values derived from oligomycin and rotenone phase.

## Comparison with OCR-Stats

In this work, we compared the maximal respiration in 26 patient cell lines to the control cell line individually as what was done in OCR-stats [@Yepez2018]. 
Overall our analysis recalled successfully all negative controls and five out of six positive controls.

Patient cell line 61818 was labeled as a positive control since this patient was found having a mutation in *SFXN4* gene.
A recent study based on erythroleukemic cell line showed that *SFXN4* knockout resulted in significant decrease in all parameters of respiration, including baseline respiration, respiratory ATP synthesis, maximal respiration, and spare respiratory capacity [@Paul2019]. 
However, our analysis showed that the maximal respiration of this patient was not significantly different from the control cell line. 
Our further analysis showed that the patient cell line only showed tenancy of having lower maximal respiration than control in one of the three repeated experiments.

We also noticed that our outcomes were not identical as what was presented in OCR-Stats [@Yepez2018]. 
Strikingly, in our analysis patient 73387 showed significant decreasing in maximal respiration compared to the control cell line. 
However, in the OCR-stats [@Yepez2018], the difference in maximal respiration was not statistically significant in this patient. 
This patient 73387 was diagnosed carrying a homozygous loss of function mutation in the *PET100* gene, which encodes a mitochondrial complex IV biogenesis factor [@Yepez2018; @Olahova2015; @Lim2014]. 
A homozygous truncating variant (c.142C>T, p.(Gln48\*)) in the *PET100* gene was found to lead to a complete loss of enzyme activity, and caused deficiency in complex IV [@Olahova2015a]. 
We think that the significant reduction of maximal respiration in patient 73387 corresponded to its deficiency in mitochondrial complex IV.

## OCRbayes is potentially used for screening mitochondrial disease

Since increasing in proton leak or a decreasing in basal or maximal respiration are indicators of mitochondrial dysfunction [@Brand2011], seahorse XF is potentially to be used for screening mitochondrial disease patients. 
An immediate question is whether we need to run a single or multiple seahorse assays for a patient. 
We suggest that whether performing a single or multiple seahorse assay depends on the effect size of perturbation. 
OCRbayes allows us to calculate posterior probability that the maximal respiration was decreased in patient even based on a single seahorse assay. 
When the perturbation effect was large, such as cell line 67375, which had the largest effect size among all the positive controls. 
The posterior probability after observing a single plate was already over 90%. 
It is reasonable to run just a single assay in this case. 
In contrast, when the perturbation effect was relatively small, for example the cell line 76065, which had the smallest effect size among all the positive controls, the posterior probability after running a single assay was modest (about 40%). 
Repeating the experiment once more increased the posterior probability from 40% to 70%. 
Thus in this case, it is beneficial to run multiple seahorse assays. 
We think that the posterior probability is an useful metric to help scientists to decide whether to perform extra seahorse assays on the patient cell lines.

## Advantages and limitations

OCRbayes has several advantages. 
The first advantage of our approach is that it incorporates various technical variations including between measurement cycle variation, between well variation as well as between plate variation during the estimation of bioenergetic measures. 
All current methods needs to choose a single data point from the three measurement cycles to represent the OCR during a particular phase. 
This procedure ignores the uncertainty and makes analysis less robust because different choice of data points may lead to different results. 
The second advantage is that OCRbayes can calculate posterior probability for difference in various bioenergetic measures based on seahorse OCR data consisting of a single plate or multiple plates. 
This is a useful feature for screening biopsy samples derived from patients with mitochondrial diseases.

One limitation of this study is that OCRbayes currently focused on human fibroblasts cell line. 
This is because our model development was restricted to the limited publicly available seahorse OCR data sets. 
In the future, OCRbayes will be extended to other cell types when more experimental data are generated.

# Reference