### 2021-03-14 code_for_prior_distribution_estimation.Rmd
prior distribution estimation

---
title: "Code for prior distribution estimation"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
source("./R/customized_functions.R")
```

# Load OCR-stats OCR data

```{r}
yepez.ocr <- read.delim("Yepez_sample_ocr_data.txt")
yepez.meta <- read.delim("Yepez_sample_meta_data.txt")
```

# Prior distributions

In the OCR-stats paper, there are in total 203 fibroblasts, in which 176 fibroblasts were measured in a single plate and all these cases have no known genetic defects. 
I will use these cases for estimating prior distributions.

```{r}
orphan_experiments <- yepez.meta %>% 
  dplyr::filter(N == 1) %>% 
  dplyr::pull(Fibroblast.id)

# find the corresponding plates
my_plates <- find_host_plate(yepez.ocr, target_cell_line = orphan_experiments) %>% 
  unique()
```

Next, find the data of the corresponding plates and keep the NHDF and target cell lines data

```{r}
my_dat <- yepez.ocr %>% 
  dplyr::filter(plate_id %in% my_plates, 
                Cell_line %in% c("NHDF", orphan_experiments))

my_dat
```

Procedures:
1. Values of three repeated measurement cycles need to be aggregated into one value
2. Replicate wells need to be aggregated and cell numbers need to be accounted
3. Difference in OCR per 1K cell between a orphan cell line and NHDF
4. Perform 1 - 3 in all 78 plates, separately

```{r}
# remove wells containing NA
wells_contain_NA <- my_dat %>% 
  dplyr::filter(is.na(OCR)) %>% 
  dplyr::select(plate_id, well) %>% 
  unique()

my_dat_filtered <- my_dat %>% 
  anti_join(wells_contain_NA)

wells_imcomplete_cycles <- my_dat_filtered %>% 
  group_by(plate_id, well, Interval) %>% 
  summarize(measurement_cycles = n()) %>% 
  dplyr::filter(measurement_cycles != 3) %>% 
  dplyr::select(plate_id, well) %>% 
  distinct()

my_dat_final <- my_dat_filtered %>% 
  anti_join(wells_imcomplete_cycles)
```

```{r}
lv0 <- my_dat_final

lv1 <- lv0 %>% 
  group_by(plate_id, Cell_line, well, cell_n, Interval) %>% 
  nest() %>% 
  mutate(lnmod = map(data, ~true_signal_measurement_error(.x$OCR))) %>% 
  tidyr::unnest(lnmod) %>% 
  dplyr::select(-data, -std.error) %>% 
  spread(term, estimate) %>% 
  ungroup()
```

First prior distribution I need is for sdlog.

```{r}
ggplot(lv1, aes(sdlog)) + 
  geom_density(aes(color = Interval)) + 
  scale_color_brewer(palette = "Dark2")

MASS::fitdistr(lv1$sdlog, densfun = "lognormal")
```

```{r}
lv2 <- lv1 %>% 
  dplyr::select(-sdlog) %>% 
  mutate(ocr_per_1k = exp(meanlog)/cell_n * 1000) %>% 
  group_by(plate_id, Cell_line, Interval) %>% 
  summarize(ocr_per_1k_true = mean(ocr_per_1k), 
            ocr_per_1k_error_log = sd(log(ocr_per_1k)))
```

The second prior distribution I need is for sigma_well

```{r}
MASS::fitdistr(lv2$ocr_per_1k_error_log, densfun = "lognormal")
```

```{r}
lv3 <- lv2 %>% 
  dplyr::filter(Cell_line == "NHDF") %>% 
  group_by(Cell_line, Interval) %>% 
  summarise(mu_ocr_per_1k_true = mean(ocr_per_1k_true), 
            sd_ocr_per_1k_true_log = sd(log(ocr_per_1k_true)))

lv3
```

The third prior distribution I need is for sigma_plate

```{r}
MASS::fitdistr(lv3$sd_ocr_per_1k_true_log, densfun = "lognormal")
```

The fourth prior distribution I need is OCR_per_1k[1] and OCR_per_1k[5]

```{r}
nhdf_basal <- lv2 %>% 
  dplyr::select(-ocr_per_1k_error_log) %>% 
  dplyr::filter(Cell_line == "NHDF", Interval == "Int1") %>% 
  spread(Interval, ocr_per_1k_true) %>% 
  dplyr::pull(Int1)

nhdf_basal

MASS::fitdistr(nhdf_basal, densfun = "lognormal")

patient_basal <- lv2 %>% 
  dplyr::select(-ocr_per_1k_error_log) %>% 
  dplyr::filter(Cell_line != "NHDF", Interval == "Int1") %>% 
  spread(Interval, ocr_per_1k_true) %>% 
  dplyr::pull(Int1)

patient_basal

MASS::fitdistr(patient_basal, densfun = "lognormal")
```

The fifth prior distribution I need is for OCR_per_1k[2] and OCR_per_1k[6]

```{r}
nhdf_oligomycin <- lv2 %>% 
  dplyr::select(-ocr_per_1k_error_log) %>% 
  dplyr::filter(Cell_line == "NHDF", Interval == "Int2") %>% 
  spread(Interval, ocr_per_1k_true) %>% 
  dplyr::pull(Int2)

nhdf_oligomycin

MASS::fitdistr(nhdf_oligomycin, densfun = "lognormal")

patient_oligomycin <- lv2 %>% 
  dplyr::select(-ocr_per_1k_error_log) %>% 
  dplyr::filter(Cell_line != "NHDF", Interval == "Int2") %>% 
  spread(Interval, ocr_per_1k_true) %>% 
  dplyr::pull(Int2)

patient_oligomycin

MASS::fitdistr(patient_oligomycin, densfun = "lognormal")
```

The sixth prior distribution I need is for OCR_per_1k[3] and OCR_per_1k[7]

```{r}
nhdf_fccp <- lv2 %>% 
  dplyr::filter(Cell_line == "NHDF", Interval == "Int3") %>% 
  dplyr::select(-ocr_per_1k_error_log) %>% 
  spread(Interval, ocr_per_1k_true) %>% 
  dplyr::pull(Int3)

nhdf_fccp

MASS::fitdistr(nhdf_fccp, densfun = "lognormal")

patient_fccp <- lv2 %>% 
  dplyr::filter(Cell_line != "NHDF", Interval == "Int3") %>% 
  dplyr::select(-ocr_per_1k_error_log) %>% 
  spread(Interval, ocr_per_1k_true) %>% 
  dplyr::pull(Int3)

patient_fccp

MASS::fitdistr(patient_fccp, densfun = "lognormal")
```

The sixth prior distribution I need is for OCR_per_1k[4] and OCR_per_1k[8]

```{r}
nhdf_rotenone <- lv2 %>% 
  dplyr::filter(Cell_line == "NHDF", Interval == "Int4") %>% 
  dplyr::select(-ocr_per_1k_error_log) %>% 
  spread(Interval, ocr_per_1k_true) %>% 
  dplyr::pull(Int4)

nhdf_rotenone

MASS::fitdistr(nhdf_rotenone, densfun = "lognormal")

patient_rotenone <- lv2 %>% 
  dplyr::filter(Cell_line != "NHDF", Interval == "Int4") %>% 
  dplyr::select(-ocr_per_1k_error_log) %>% 
  spread(Interval, ocr_per_1k_true) %>% 
  dplyr::pull(Int4)

patient_rotenone

MASS::fitdistr(patient_rotenone, densfun = "lognormal")
```